---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


### Adrian Trujillo


```{r echo=FALSE , results='hide'}
library(knitr)
suppressMessages(library (dplyr))
library(lattice)
library(ggplot2)
opts_chunk$set(echo = TRUE, fig.height=5, fig.width=5, fig.path="figures/")
```



## 1 Code for reading in the dataset and/or processing the data
```{r}
activity <- read.csv("activity.csv", header = TRUE, sep = ',', colClasses = c("numeric", "character", "integer"), stringsAsFactors = F)
str(activity)
```


## 2 Histogram of the total number of steps taken each day
```{r}
complete_days_only <- activity[complete.cases(activity), ]
step_summary  <-  complete_days_only %>% 
                  group_by(date) %>% 
                  summarise(daily_step_count = sum(steps),.groups = 'drop')

hist(step_summary$daily_step_count, 
    main = "Histogram of total steps per day",
    xlab = "Range of step totals",
    ylab = "Number of totals in range",
    border = "blue",
    col = heat.colors(4),
    las = 1,
    ylim = c(0, 30))
```


## 3 Mean and median number of steps taken each day
```{r }
mean_steps <- mean(step_summary$daily_step_count, na.rm = TRUE)
median_steps <- median(step_summary$daily_step_count, na.rm = TRUE)
mean_and_median <- data.frame(Mean = mean_steps, Median = median_steps)
mean_and_median
```


## 4 Time series plot of the average number of steps taken
```{r }
x  <- complete_days_only %>% 
      group_by(interval) %>% 
      summarise(avg_interval = mean(steps),.groups = 'drop')

plot(x$interval, 
     x$avg_interval, 
     type = "l", 
     las = 1, 
     col = "brown", 
     main = "Average Steps within Intervals",
     lwd = 2,
     col.main = "blue",
     font.main = 4,
     xlab = "Daily Intervals",
     ylab = "Step Averages"
     )
```

## 5 The 5-minute interval that, on average, contains the maximum number of steps
```{r }
x[which.max(x$avg_interval), ]

```




## 6 Code to describe and show a strategy for imputing missing data
Total number of rows:
```{r }
nrow(activity)
```

Rows without NA

```{r }
nrow(complete_days_only)
```
Rows with NA

```{r }
difference <- nrow(activity) - nrow(complete_days_only)
difference
```

Fill the missing data with ramdom numbers
```{r }
set.seed(10)
z  <- floor(runif(nrow(activity), 
                  min = min(activity$steps, na.rm = T), 
                  max = max(activity$steps, na.rm = T)/10))
w <- which(is.na(activity$steps))
activity$steps[w] <- z[w]
```

## 7 Histogram of the total number of steps taken each day after missing values are imputed
```{r }
complete_data  <- activity %>% 
                  group_by(date) %>% 
                  summarise(daily_step_count = sum(steps),.groups = 'drop')

hist(complete_data$daily_step_count, 
    breaks = 10,
    main = "Histogram of total steps per day",
    xlab = "Range of step totals",
    ylab = "Number of totals in range",
    border = "green",
    col = heat.colors(12),
    las = 1,
    ylim = c(0, 25))
```

## 8 Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
Determine if there are differences in activity patterns between weekdays and weekends
```{r }
activity$date <- as.POSIXct(activity$date)
activity$dayType  <- ifelse(weekdays(activity$date) %in% c(weekdays(as.POSIXct("2020-06-06")),weekdays(as.POSIXct("2020-06-06"))), "weekend", "weekday")

activity$dayType  <- as.factor(activity$dayType)
q  <- activity %>% 
      group_by(dayType, interval) %>% 
      summarise(daily_step_count = sum(steps),.groups = 'drop')
q
```

Create a time series plot of total steps within intervals comparing weekdays vs weekends
```{r }
with(q, 
      xyplot(daily_step_count ~ interval | dayType, 
      type = "l",      
      main = "Total Steps within Intervals",
      xlab = "Daily Intervals",
      ylab = "Total Steps"))
```

## The diagram above suggests that people walk less on weekends


